---
title: "Continuous Effects of Early Adversities with Interactions Considered"
author: "Jiajun"
date: "03/05/2021"
output:
  rmdformats::readthedown:
    css: ["data/css/text_style.css", "data/css/table_style.css"]
    theme: sandstone
    toc_depth: 3
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Library and Data

```{r load-library-and-data, class.source='fold-show'}
library(survival)
library(glmnet)
library(kableExtra)
library(tidyverse)
library(GGally)
library(penalized)

raw_data <- read_csv("data/Aim3-early-adversity-subjects-F-survival-2021-02-17.csv")
raw_data_sub <- raw_data %>%
  dplyr::select(dead, age,
                rain, density, mom_rank, mom_sci,
                adv_mom, adv_sib) %>%
  mutate(adv_mom = as.numeric(adv_mom),
         adv_sib = as.numeric(adv_sib)) %>%
  filter(age > 0)
```

```{r covariates-info-table}
covariate_names <- c("adv_mom", "adv_sib",
                     "rain", "density", "mom_rank", "mom_sci")
covariate_types <- c(rep("Logical", 2), rep("Numeric", 4))
covariate_meanings <- c("Mom dies before age 4",
                       "A close sibling is born",
                       "Group size at birth",
                       "Rainfall in the first year of life",
                       "Mom rank",
                       "Mom Social connection")
covariate_summary <- c("139/562",
                       "149/562",
                       "340.5(134.7)",
                       "26.6(8.6)",
                       "0.5(0.3)",
                       "0.1(0.75)")

data.frame(
  list(types = covariate_types,
       meanings = covariate_meanings,
       summary = covariate_summary)
) %>%
  `row.names<-`(covariate_names) %>%
  kable(col.names = c("Type", "Description", "Summary")) %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 18)
```


```{r}
ggpairs(data = raw_data_sub %>%
          select(- c(dead, age)) %>%
          mutate(adv_mom = as.factor(adv_mom),
                 adv_sib = as.factor(adv_sib)),
        lower = list(continuous = wrap("points", size = 0.5, alpha = 0.5),
                     combo = wrap("facethist", bins = 20))) +
  theme_classic()
```


<br>

## Process

```{r create-dataset-subgroups, echo=FALSE}
age_below_1 <- raw_data_sub %>%
  mutate(dead = if_else(age > 1, FALSE, dead)) %>%
  mutate(age = if_else(age > 1, 1, age))

age_above_1_below_5 <- raw_data_sub %>%
  filter(age >= 1) %>%
  mutate(dead = if_else(age > 5, FALSE, dead)) %>%
  mutate(age = if_else(age > 5, 5, age))

age_above_5 <- raw_data_sub %>%
  filter(age >= 5)
```

- As before, the dataset is divided into infant, juvenile, and adults phase based on the cut of `age` at 1 and 5.
- For each subgroup
  - First run a simple regression with one variable at a time. (e.g. survival ~ rain + (maybe rain^2)).
  - Then put all variables in one model.
  - Lastly, include the interactions which is significant when modeled individually, apply variable selection using lasso, and run the model on picked variables.


<br>

## Age below 1 {.tabset .tabset-fade .tabset-pills}

### adv_mom

```{r}
g1_adv_mom_coxph <- coxph(Surv(age, dead) ~ adv_mom, data = age_below_1)
g1_adv_mom_coxph
```

```{r echo=FALSE}
termplot(g1_adv_mom_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### rain

```{r}
g1_rain_coxph <- coxph(Surv(age, dead) ~ pspline(rain, df = 4), data = age_below_1)
g1_rain_coxph
```

```{r echo=FALSE}
termplot(g1_rain_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### density

```{r}
g1_density_coxph <- coxph(Surv(age, dead) ~ pspline(density, df = 4), data = age_below_1)
g1_density_coxph
```

```{r echo=FALSE}
termplot(g1_density_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### mom_rank

```{r}
g1_momrank_coxph <- coxph(Surv(age, dead) ~ pspline(mom_rank, df = 4), data = age_below_1)
g1_momrank_coxph
```

```{r echo=FALSE}
termplot(g1_momrank_coxph, se=T, terms=1, ylabs = "Log hazard")
```


### mom_sci

```{r}
g1_momsci_coxph <- coxph(Surv(age, dead) ~ pspline(mom_sci, df = 4), data = age_below_1)
g1_momsci_coxph
```

```{r echo=FALSE}
termplot(g1_momsci_coxph, se=T, terms=1, ylabs = "Log hazard")
```


## {-}

## Age below one: All in one model {.tabset .tabset-fade .tabset-pills}

### Main effects

```{r}
g1_all <- coxph(
  Surv(age, dead) ~ adv_mom + rain + density + mom_rank + mom_sci,
  data = age_below_1
)
print(g1_all)
```


### Lasso path

```{r}
penalized_covariates <- model.matrix(
  ~ adv_mom + rain + density + mom_rank + mom_sci +
    rain:mom_sci + adv_mom:rain + adv_mom:mom_sci,
  data = age_below_1
) %>% .[, colnames(.) != "(Intercept)"]

opt <- optL1(
  Surv(age_below_1$age, age_below_1$dead),
  penalized = penalized_covariates,
  standardize = T,
  fold = 10,
  trace = F
)

stepwise <- penalized(
  Surv(age_below_1$age, age_below_1$dead),
  penalized = penalized_covariates,
  standardize = T,
  steps = 20,
  trace = F
)

plotpath(
  stepwise,
  labelsize = 0.8,
  standardize = T
)

abline(
  v = opt$lambda,
  col = "gray",
  lwd = 2
)
```

### Results with picked predictors

```{r}
g1_opt_lasso <- penalized(
  Surv(age_below_1$age, age_below_1$dead),
  penalized = penalized_covariates,
  standardize = T,
  lambda1 = opt$lambda,
  trace = F
)

nonzero_varnames <- names(g1_opt_lasso@penalized)[g1_opt_lasso@penalized != 0]
formula <- paste0(" Surv(age, dead) ~ ", paste(nonzero_varnames, collapse = " + "))
g1_final_coxph <- coxph(
  as.formula(formula),
  data = age_below_1
)

g1_final_coxph
```



## {-}

<br><br>

## Age above 1 and below 5 {.tabset .tabset-fade .tabset-pills}

### adv_mom

```{r}
g2_adv_mom_coxph <- coxph(Surv(age, dead) ~ adv_mom, data = age_above_1_below_5)
g2_adv_mom_coxph
```

```{r echo=FALSE}
termplot(g2_adv_mom_coxph, se=T, terms=1, ylabs = "Log hazard")
```


### adv_sib

```{r}
g2_adv_sib_coxph <- coxph(Surv(age, dead) ~ adv_sib, data = age_above_1_below_5)
g2_adv_sib_coxph
```

```{r, echo=FALSE}
termplot(g2_adv_sib_coxph, se=T, terms=1, ylabs = "Log hazard")
```


### rain

```{r}
g2_rain_coxph <- coxph(Surv(age, dead) ~ pspline(rain, df = 4), data = age_above_1_below_5)
g2_rain_coxph
```

```{r echo=FALSE}
termplot(g2_rain_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### density

```{r}
g2_density_coxph <- coxph(Surv(age, dead) ~ pspline(density, df = 4), data = age_above_1_below_5)
g2_density_coxph
```

```{r echo=FALSE}
termplot(g2_density_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### mom_rank

```{r}
g2_momrank_coxph <- coxph(Surv(age, dead) ~ pspline(mom_rank, df = 4), data = age_above_1_below_5)
g2_momrank_coxph
```

```{r echo=FALSE}
termplot(g2_momrank_coxph, se=T, terms=1, ylabs = "Log hazard")
```


### mom_sci

```{r}
g2_momsci_coxph <- coxph(Surv(age, dead) ~ pspline(mom_sci, df = 4), data = age_above_1_below_5)
g2_momsci_coxph
```

```{r echo=FALSE}
termplot(g2_momsci_coxph, se=T, terms=1, ylabs = "Log hazard")
```


## {-}

## Age above 1 below 5: All in one model {.tabset .tabset-fade .tabset-pills}

### Main effects

```{r}
g2_all <- coxph(
  Surv(age, dead) ~ adv_mom + adv_sib + rain + density + mom_rank + mom_sci,
  data = age_above_1_below_5
)
print(g2_all)
```


### Lasso path

```{r}
penalized_covariates <- model.matrix(
  ~adv_mom + adv_sib + rain + density + I(density^2) + mom_rank + mom_sci +
    adv_mom:density + adv_mom:I(density^2),
  data = age_above_1_below_5
) %>% .[, colnames(.) != "(Intercept)"]

opt <- optL1(
  Surv(age_above_1_below_5$age, age_above_1_below_5$dead),
  penalized = penalized_covariates,
  standardize = T,
  fold = 10,
  trace = F
)

stepwise <- penalized(
  Surv(age_above_1_below_5$age, age_above_1_below_5$dead),
  penalized = penalized_covariates,
  standardize = T,
  steps = 20,
  trace = F
)

plotpath(
  stepwise,
  labelsize = 0.8,
  standardize = T
)

abline(
  v = opt$lambda,
  col = "gray",
  lwd = 2
)
```

### Results with picked predictors

```{r}
g2_opt_lasso <- penalized(
  Surv(age_above_1_below_5$age, age_above_1_below_5$dead),
  penalized = penalized_covariates,
  standardize = T,
  lambda1 = opt$lambda,
  trace = F
)

nonzero_varnames <- names(g2_opt_lasso@penalized)[g2_opt_lasso@penalized != 0]
formula <- paste0(" Surv(age, dead) ~ ", paste(nonzero_varnames, collapse = " + "))
g2_final_coxph <- coxph(
  as.formula(formula),
  data = age_above_1_below_5
)

g2_final_coxph
```


## {-}

<br><br>

## Age above 5 {.tabset .tabset-fade .tabset-pills}

### adv_mom

```{r}
g3_adv_mom_coxph <- coxph(Surv(age, dead) ~ adv_mom, data = age_above_5)
g3_adv_mom_coxph
```

```{r echo=FALSE}
termplot(g3_adv_mom_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### adv_sib

```{r}
g3_adv_sib_coxph <- coxph(Surv(age, dead) ~ adv_sib, data = age_above_5)
g3_adv_sib_coxph
```

```{r, echo=FALSE}
termplot(g3_adv_sib_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### rain

```{r}
g3_rain_coxph <- coxph(Surv(age, dead) ~ pspline(rain, df = 4), data = age_above_5)
g3_rain_coxph
```

```{r echo=FALSE}
termplot(g3_rain_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### density

```{r}
g3_density_coxph <- coxph(Surv(age, dead) ~ pspline(density, df = 4), data = age_above_5)
g3_density_coxph
```

```{r echo=FALSE}
termplot(g3_density_coxph, se=T, terms=1, ylabs = "Log hazard")
```

### mom_rank

```{r}
g3_momrank_coxph <- coxph(Surv(age, dead) ~ pspline(mom_rank, df = 4), data = age_above_5)
g3_momrank_coxph
```

```{r echo=FALSE}
termplot(g3_momrank_coxph, se=T, terms=1, ylabs = "Log hazard")
```


### mom_sci

```{r}
g3_momsci_coxph <- coxph(Surv(age, dead) ~ pspline(mom_sci, df = 4), data = age_above_5)
g3_momsci_coxph
```

```{r echo=FALSE}
termplot(g3_momsci_coxph, se=T, terms=1, ylabs = "Log hazard")
```


## {-}

## Age above 5: All in one model {.tabset .tabset-fade .tabset-pills}

### Main effects

```{r}
g3_all <- coxph(
  Surv(age, dead) ~ adv_mom + adv_sib + rain + density + mom_rank + mom_sci,
  data = age_above_5
)
print(g3_all)
```


### Lasso path

```{r}
penalized_covariates <- model.matrix(
  ~adv_mom + adv_sib + rain + density + I(density^2) + mom_rank + mom_sci +
    adv_mom:density + adv_mom:I(density^2),
  data = age_above_5
) %>% .[, colnames(.) != "(Intercept)"]

opt <- optL1(
  Surv(age_above_5$age, age_above_5$dead),
  penalized = penalized_covariates,
  standardize = T,
  fold = 10,
  trace = F
)

stepwise <- penalized(
  Surv(age_above_5$age, age_above_5$dead),
  penalized = penalized_covariates,
  standardize = T,
  steps = 20,
  trace = F
)

plotpath(
  stepwise,
  labelsize = 0.8,
  standardize = T
)

abline(
  v = opt$lambda,
  col = "gray",
  lwd = 2
)
```

### Results with picked predictors

```{r}
g3_opt_lasso <- penalized(
  Surv(age_above_5$age, age_above_5$dead),
  penalized = penalized_covariates,
  standardize = T,
  lambda1 = opt$lambda,
  trace = F
)

nonzero_varnames <- names(g3_opt_lasso@penalized)[g3_opt_lasso@penalized != 0]
formula <- paste0(" Surv(age, dead) ~ ", paste(nonzero_varnames, collapse = " + "))
g3_final_coxph <- coxph(
  as.formula(formula),
  data = age_above_5
)

g3_final_coxph
```


## {-}

<br><br>

## Discussion

- Few of `rain`, `density`, `mom_rank`, and `mom_sci` have shown significance in the cox model on the three datasets, either as a single variable or put together.
- For the infant group(age < 1), `rain` is significant with p-value 0.03. However, no significance is shown for the nonlinear part.
- Nonlinear part of `density` is significant for the juvenile(1 < age < 5) and adult(age > 5) group, with p-value 0.023 and 0.03 respectively.




